---
title: 'Coursera Data Scientist Track - Statistical Inference Course Project, Part One: Simulation Exercises'
author: "Pieter Jonckheere"
output:
  pdf_document:
    fig_height: 4
  html_document: default
---

In this project we will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. We will investigate the distribution of averages of 40 exponentials. We will do this based on a thousand simulations.


Perform a thousand simulated averages of 40 exponentials.

```{r}
set.seed(3)
lambda <- 0.2
number_sim <- 1000
sample_size <- 40
sim <- matrix(rexp(number_sim*sample_size, rate=lambda), number_sim, sample_size)
row_means <- rowMeans(sim)
```

The distribution of sample means is as follows.

```{r echo=FALSE}
# plot the histogram of averages
hist(row_means, breaks=50, prob=TRUE,
     main="Distribution of averages of samples,
     drawn from exponential distribution (lambda = 0.2)",
     xlab="")
# density of the averages of samples
lines(density(row_means))
# theoretical center of distribution
abline(v=1/lambda, col="green")
# theoretical density of the averages of samples
xfit <- seq(min(row_means), max(row_means), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(sample_size)))
lines(xfit, yfit, pch=22, col="green", lty=2)
# add legend
legend('topright', c("simulation", "theoretical"), lty=c(1,2), col=c("black", "green"))
```

The distribution of sample means is centered at `r mean(row_means)` and the theoretical center of the 
distribution is $\lambda^{-1}$ = `r 1/lambda`.
The variance of sample means is `r var(row_means)`, where the theoretical variance
of the distribution is $\sigma^2 / n = 1/(\lambda^2 n) = 1/(0.04 \times 40)$ =
`r 1/(0.04 * 40)`.

Due to the central limit theorem, the averages of samples follow a normal
distribution. The figure above also shows the density computed using the histogram and the
normal density plotted with theoretical mean and variance values. 


